Capturing Subdiffusive Solute Dynamics and Predicting Selectivity in Nanoscale Pores with Time Series Modeling

### 2020-01-21 Ben_Manuscripts/stochastic_transport/Final_Draft.tex
Minor changes

  Mathematically modeling complex transport phenomena at the molecular level 
  can be a powerful tool for identifying transport mechanisms and predicting
  macroscopic properties. We use two different stochastic time series models,
  parameterized from long molecular dynamics (MD) simulation trajectories of
  a cross-linked H\textsubscript{II} phase lyotropic liquid crystal (LLC) 
  membrane, in order to predict solute mean squared displacements (MSDs) 
  and solute flux, and thus solute selectivity, in macroscopic length pores. 
  First, using anomalous diffusion theory, we show how solute dynamics can be
  modeled as a fractional diffusion process subordinate to a continuous time 
  random walk. From the MD simulations, we parameterize the distribution of 
  dwell times, hop lengths between dwells and correlation between hops. We 
  explore two variations of the anomalous diffusion model. The first variation
  applies a single set of parameters to the solute displacements and the second
  applies two sets of parameters based on the solute's radial distance from the
  closest pore center. Next, we generalize Markov state models, treating the
  configurational state of the system as a Markov process, each
  of which has separate transport properties. For each state and transition
  between states, we parameterize the distribution and temporal correlation 
  structure of positional fluctuations as a means of characterization and to 
  allow us to predict solute MSDs. We show that both models reasonably reproduce
  the MSDs calculated from MD simulations. 
  However, qualitative differences between MD and 
  %simulated MSDDM 
  MSDDM-generated  %MRS: abbreviation not defined first? Maybe eliminate here, just call it the Markov State dependent model
  trajectories
  may limit its usefulness. 
  Finally, we demonstrate how one can use 
  %our 
  these 
  models to estimate flux of a solute
  across a macroscopic-length pore and, based on those quantities, the membrane's selectivity towards each solute.
  This work helps to connect microscopic chemically-dependent solute 
  motions 
  %MRS8:  Added the below, since for simple diffusion, you just calculate the diffusion constant. 
  %and 
  that do not follow simple diffusive behavior with 
  macroscopic membrane performance.

>>>>

  Mathematically modeling complex transport phenomena at the molecular level 
  can be a powerful tool for identifying transport mechanisms and predicting
  macroscopic properties. We use two different stochastic time series models,
  parameterized from long molecular dynamics (MD) simulation trajectories of
  a cross-linked H\textsubscript{II} phase lyotropic liquid crystal (LLC) 
  membrane, in order to predict solute mean squared displacements (MSDs) 
  and solute flux, and thus solute selectivity, in macroscopic length pores. 
  First, using anomalous diffusion theory, we show how solute dynamics can be
  modeled as a fractional diffusion process subordinate to a continuous time 
  random walk. From the MD simulations, we parameterize the distribution of 
  dwell times, hop lengths between dwells and correlation between hops. We 
  explore two variations of the anomalous diffusion model. The first variation
  applies a single set of parameters to the solute displacements and the second
  applies two sets of parameters based on the solute's radial distance from the
  closest pore center. Next, we generalize Markov state models, treating the
  configurational state of the system as a Markov process, each
  of which has separate transport properties. For each state and transition
  between states, we parameterize the distribution and temporal correlation 
  structure of positional fluctuations as a means of characterization and to 
  allow us to predict solute MSDs. We show that both models reasonably reproduce
  the MSDs calculated from MD simulations. However, qualitative differences 
  between MD and Markov state dependent model-generated trajectories may limit 
  its usefulness. Finally, we demonstrate how one can use these models to 
  estimate flux of a solute across a macroscopic-length pore and, based on 
  those quantities, the membrane's selectivity towards each solute. This work 
  helps to connect microscopic chemically-dependent solute motions that do not
  follow simple diffusive behavior with macroscopic membrane performance.
